>

### 2024-07-16 docs/abstract/_abstract-eupha.qmd
Initial draft for ISPOR 2024

**Methods:**
Exploratory time-series analysis was performed in a static cohort from IADB.nl with at least one prescription of psychopharmaca from 2018 to 2022. The prescription was evaluated daily using anatomical therapeutic chemical (ATC) classification. A drug prescription network was constructed to capture dose-adjusted co-prescription. Eigenvector centrality ($c_e$) was computed as a measure of relative nodal importance, which reflects the degree of medication concurrence in a network. The number of medication claims ($n_c$) and eigenvector centrality were extracted as weekly-aggregated time-series data.

@@

**Conclusion:**
Seven medications have a high concurrence on a population level. Positive time trends in both eigenvector centrality and claims suggest that increasing uses of antidepressants are related to co-prescription with other medication classes. We recommend constructing a medication claim network to evaluate periodic co-prescription as a public health monitoring procedure.

>>>>

**Methods:**
Exploratory time-series analysis was performed in a closed cohort from IADB.nl with at least one prescription of anxiolytics/antidepressants from 2018 to 2022. The prescription was evaluated daily using anatomical therapeutic chemical (ATC) classification. A drug prescription network was constructed to capture dose-adjusted co-prescription. Eigenvector centrality ($c_e$) was computed as a measure of relative nodal importance, which reflects the degree of medication concurrence in a network. The number of medication claims ($n_c$) and eigenvector centrality were extracted as weekly-aggregated time-series data.

@@

**Conclusion:**
Seven medications, three of which are psychopharmaca, had a high eigenvector centrality, indicating concurrence with other medication groups on a population level. Positive time trends in both eigenvector centrality and claims suggest that increasing uses of antidepressants was related to co-prescription with other medication classes. Constructing a medication claim network enriches the exploratory analysis of medication use at the population level.

### 2024-07-16 docs/abstract/_abstract-full.qmd
Initial draft for ISPOR 2024

**Methods:**
Exploratory time-series analysis was performed using the University of Groningen prescription database, IADB.nl. A static cohort with at least one prescription of psychopharmaca from 2018 to 2022 was extracted. The prescription was evaluated daily using anatomical therapeutic chemical (ATC) classification, where psychopharmaca was evaluated at ATC level 3, drugs for the nervous system at ATC level 2, and other medications at ATC level 1, distinguished into a total of 24 medication classes. A drug prescription network was constructed as an undirected graph to capture dose-adjusted co-prescription. Eigenvector centrality ($c_e$) was computed as a measure of relative nodal importance, which reflects the degree of medication concurrence in a network. The number of medication claims ($n_c$), number of patients, claim-to-patient ratio, and eigenvector centrality were extracted as weekly-aggregated time-series data. Singular spectrum analysis was performed to decompose the data into its constituting trend, periodicity, and white noises.

**Results:**
Antidepressants (ATC `N06A`, $c_e$: `r N06A$ce`, $n_c$: `r N06A$nc`) and anxiolytics (`N05B`, $c_e$: `r N05B$ce`, $n_c$: `r N05B$nc`) are psychopharmaca with high eigenvector centrality. These classes of medication are likely to be co-prescribed alongside other medications. Further groups with high eigenvector centrality were medications for alimentary and metabolism (`A01-A16`, $c_e$: `r A0x$ce`, $n_c$: `r A0x$nc`), blood (`B01-B06`, $c_e$: `r B0x$ce`, $n_c$: `r B0x$nc`), cardiovascular (`C01-C10`, $c_e$: `r C0x$ce`, $n_c$: `r C0x$nc`), analgesics (`N02`, $c_e$: `r N02$ce`, $n_c$: `r N02$nc`), and respiratory (`R01-R07`, $c_e$: `r R0x$ce`, $n_c$: `r R0x$nc`). Among eight medications with high eigenvector centrality, only antidepressants had a positive time trend for both eigenvector centrality and actual number of claims.

**Discussion:**
Seven medications, three of which are psychopharmaca, have a high eigenvector centrality, implying its concurrence on a population level. Positive time trends in both eigenvector centrality and claims suggest that increasing uses of antidepressants are related to co-prescription with other medication classes. Constructing a medication claim network enriches the exploratory analysis of medication use at the population level.

>>>>

**Methods:**
Exploratory time-series analysis was performed using the University of Groningen prescription database, IADB.nl. A closed cohort with at least one prescription of anxiolytics/antidepressants from 2018 to 2022 was extracted. The prescription was evaluated daily using anatomical therapeutic chemical (ATC) classification, where psychopharmaca was evaluated at ATC level 4, drugs for the nervous system at ATC level 3, and other medications at ATC level 1, leading to a total of 24 medication classes. A drug prescription network was constructed as an undirected graph to capture dose-adjusted co-prescription. Eigenvector centrality ($c_e$) was computed as a measure of relative nodal importance, which reflects the degree of medication concurrence in a network. The number of medication claims ($n_c$), number of patients, claim-to-patient ratio, and eigenvector centrality were extracted as weekly-aggregated time-series data. Singular spectrum analysis was performed to decompose the data into its constituting trend, periodicity, and white noises.

**Results:**
Antidepressants (ATC `N06A`, $c_e$: `r N06A$ce`, $n_c$: `r N06A$nc`) and anxiolytics (`N05B`, $c_e$: `r N05B$ce`, $n_c$: `r N05B$nc`) are psychopharmaca with high eigenvector centrality. These classes of medication are likely to be co-prescribed alongside other medications. Further groups with high eigenvector centrality were medications for alimentary and metabolism (`A01-A16`, $c_e$: `r A0x$ce`, $n_c$: `r A0x$nc`), blood (`B01-B06`, $c_e$: `r B0x$ce`, $n_c$: `r B0x$nc`), cardiovascular (`C01-C10`, $c_e$: `r C0x$ce`, $n_c$: `r C0x$nc`), analgesics (`N02`, $c_e$: `r N02$ce`, $n_c$: `r N02$nc`), and respiratory (`R01-R07`, $c_e$: `r R0x$ce`, $n_c$: `r R0x$nc`). Among the seven medication groups with high eigenvector centrality, only antidepressants showed a positive time trend for both eigenvector centrality and actual number of claims.

**Discussion:**
Seven medications, three of which are psychopharmaca, had a high eigenvector centrality, indicating concurrence with other medication groups on a population level. Positive time trends in both eigenvector centrality and claims suggest that increasing uses of antidepressants was related to co-prescription with other medication classes. Constructing a medication claim network enriches the exploratory analysis of medication use at the population level.

### 2024-07-16 docs/abstract/_abstract-ispor.qmd
Initial draft for ISPOR 2024

**Methods:**
Exploratory time-series analysis was performed using the University of Groningen prescription database, IADB.nl. A static cohort with at least one prescription of psychopharmaca from 2018 to 2022 was extracted. The prescription was evaluated daily using anatomical therapeutic chemical (ATC) classification, where psychopharmaca was evaluated at ATC level 3, drugs for the nervous system at ATC level 2, and other medications at ATC level 1, distinguished into a total of 24 medication classes. A drug prescription network was constructed as an undirected graph to capture dose-adjusted co-prescription. Eigenvector centrality ($c_e$) was computed as a measure of relative nodal importance, which reflects the degree of medication concurrence in a network. The number of medication claims ($n_c$), number of patients, claim-to-patient ratio, and eigenvector centrality were extracted as weekly-aggregated time-series data. Singular spectrum analysis was performed to decompose the data into its constituting trend, periodicity, and white noises.

**Results:**
Antidepressants (ATC `N06A`, $c_e$: `r N06A$ce`, $n_c$: `r N06A$nc`) and anxiolytics (`N05B`, $c_e$: `r N05B$ce`, $n_c$: `r N05B$nc`) are psychopharmaca with high eigenvector centrality. These classes of medication are likely to be co-prescribed alongside other medications. Further groups with high eigenvector centrality were medications for alimentary and metabolism (`A01-A16`, $c_e$: `r A0x$ce`, $n_c$: `r A0x$nc`), blood (`B01-B06`, $c_e$: `r B0x$ce`, $n_c$: `r B0x$nc`), cardiovascular (`C01-C10`, $c_e$: `r C0x$ce`, $n_c$: `r C0x$nc`), analgesics (`N02`, $c_e$: `r N02$ce`, $n_c$: `r N02$nc`), and respiratory (`R01-R07`, $c_e$: `r R0x$ce`, $n_c$: `r R0x$nc`). Among eight medications with high eigenvector centrality, only antidepressants had a positive time trend for both eigenvector centrality and actual number of claims.

**Discussion:**
Seven medications, three of which are psychopharmaca, have a high eigenvector centrality, implying its concurrence on a population level. Positive time trends in both eigenvector centrality and claims suggest that increasing uses of antidepressants are related to co-prescription with other medication classes. Constructing a medication claim network enriches the exploratory analysis of medication use at the population level.

>>>>

**Methods:**
Exploratory time-series analysis was performed using the University of Groningen prescription database, IADB.nl. A closed cohort with at least one prescription of anxiolytics/antidepressants from 2018 to 2022 was extracted. The prescription was evaluated daily using anatomical therapeutic chemical (ATC) classification, where psychopharmaca was evaluated at ATC level 4, drugs for the nervous system at ATC level 3, and other medications at ATC level 1, distinguished into a total of 24 medication classes. A drug prescription network was constructed as an undirected graph to capture dose-adjusted co-prescription. Eigenvector centrality ($c_e$) was computed as a measure of relative nodal importance, which reflects the degree of medication concurrence in a network. The number of medication claims ($n_c$), number of patients, claim-to-patient ratio, and eigenvector centrality were extracted as weekly-aggregated time-series data. Singular spectrum analysis was performed to decompose the data into its constituting trend, periodicity, and white noises.

**Results:**
Antidepressants (ATC `N06A`, $c_e$: `r N06A$ce`, $n_c$: `r N06A$nc`) and anxiolytics (`N05B`, $c_e$: `r N05B$ce`, $n_c$: `r N05B$nc`) are psychopharmaca with high eigenvector centrality. These classes of medication are likely to be co-prescribed alongside other medications. Further groups with high eigenvector centrality were medications for alimentary and metabolism (`A01-A16`, $c_e$: `r A0x$ce`, $n_c$: `r A0x$nc`), blood (`B01-B06`, $c_e$: `r B0x$ce`, $n_c$: `r B0x$nc`), cardiovascular (`C01-C10`, $c_e$: `r C0x$ce`, $n_c$: `r C0x$nc`), analgesics (`N02`, $c_e$: `r N02$ce`, $n_c$: `r N02$nc`), and respiratory (`R01-R07`, $c_e$: `r R0x$ce`, $n_c$: `r R0x$nc`). Among the seven medication groups with high eigenvector centrality, only antidepressants showed a positive time trend for both eigenvector centrality and actual number of claims.

**Discussion:**
Seven medications, three of which are psychopharmaca, had a high eigenvector centrality, indicating concurrence with other medication groups on a population level. Positive time trends in both eigenvector centrality and claims suggest that increasing uses of antidepressants was related to co-prescription with other medication classes. Constructing a medication claim network enriches the exploratory analysis of medication use at the population level.

### 2024-07-16 docs/article.qmd
Initial draft for ISPOR 2024

---
title: >
  Eigenvector centrality as a metric of polypharmacy signifies antidepressants
  as a central hub of the drug prescription network

@@

Surveilling polypharmacy regimens requires a holistic approach since drug prescription highly depends on the physician's prescribing habits [@Stassen2021], patient-physician relationship, and pharmaceutical market, which all bear an interplay with the decision-making process [@Cavallo2012]. A holistic polypharmacy surveillance system needs to describe the importance of polypharmacy by incorporating both qualitative and quantitative measures [@Sirois2016]. Quantitatively attributing importance to each medication is challenging, accounting for polypharmacy as a complex and interlinked medication uses. To reduce the complexity, graph theory offers a complementary approach by analyzing polypharmacy as a drug prescription network (DPN). Transforming prescription registry into a DPN has been previously investigated, where DPN-based approach represented the general trend in co-prescription [@Cavallo2012; @Bazzoni2015].

>>>>

---
title: >
  A temporal network analysis of drug co-prescription around antidepressants and anxiolytics uses

@@

Surveilling polypharmacy regimens requires a holistic approach since drug prescription highly depends on the physician's prescribing habits [@Shrivastava2013], patient-physician relationship, and pharmaceutical market, which all bear an interplay with the decision-making process [@Cavallo2012]. A holistic polypharmacy surveillance system needs to describe the importance of polypharmacy by incorporating both qualitative and quantitative measures [@Sirois2016]. Quantitatively attributing importance to each medication is challenging, accounting for polypharmacy as a complex and interlinked medication uses. To reduce the complexity, graph theory offers a complementary approach by analyzing polypharmacy as a drug prescription network (DPN). Transforming prescription registry into a DPN has been previously investigated, where DPN-based approach represented the general trend in co-prescription [@Cavallo2012; @Bazzoni2015].
